\chapter{Acknowledgements}
\begin{paracol}{2}
    \small
    First and foremost, I would like to thank my supervisor, who has been a solid and constant---should I say iner\-tial?---frame of reference since my very first day at university. His guidance, patience, and ability to steer me in the right direction have been invaluable during this journey. I really enjoyed working on the project he proposed, as it gave me a deeper insight on some topics of modern physics and the chance to work with some of the tools of theoretical physics.
    \switchcolumn
    Prima di tutto, vorrei ringraziare il mio relatore, che fin dall'primo istante del primo anno è stato un solido e costante sistema di riferimento inerziale. La sua competenza e disponibilità, ma anche la sua pazienza, sono state inestimabili durante il mio percorso. Ho apprezzato il progetto di tesi che mi ha proposto in quanto mi ha offerto nuovi sguardi sulla fisica moderna e mi ha permesso di confrontarmi con alcuni strumenti della fisica teorica.
    \switchcolumn*
    A heartfelt thanks goes to my senior colleague Santo, living up to his name in every possible sense. His generosity with both his time and his experience---and his beautifully typeset lecture notes---allowed me to pass more exams than I care to admit. Without him I would certainly still be in my third second year, using textbooks to wipe my tears.
    \switchcolumn
    Un ringraziamento speciale va al mio collega più grande Santo, di nome e di fatto. La sua gentilezza e il suo supporto, il suo tempo, la sua esperienza e i suoi meravigliosi appunti in \LaTeX\ mi hanno aiutato a passare più esami di quel che tendo ad ammettere. Senza il suo aiuto probabilmente sarei ancora al terzo secondo anno ad asciugarmi le lacrime coi libri.
    \switchcolumn*
    I am also deeply grateful to my dearest friends, who turned this academic marathon into something more human. To my colleagues Martina and Arianna, exceptional desk mates and fellow partners in breakdowns for the past few years. To Gabriele, with whom I studied endless nights until 4 in the morning: our coffee consumption should probably be considered a separate research project. To my best friend Porto, who has been like a brother to me, always putting up with my whining and also my main income in terms of laughters and advice. And naturally, to all of my other dear friends, like \emph{the Bibbins} and \emph{the Penguins} who I would love to mention one by one if it weren't for the fact that it's 3:40 in the morning as I write, trying my best not to leave anyone out.
    \switchcolumn
    Sono profondamente grato anche ai miei amici più cari, che hanno reso la maratona universitaria un po' più sopportabile. Alle mie colleghe Martina e Arianna, fantastiche compagne di banco e di sclero per gli ultimi anni. A Gabriele, con cui ho studiato innumerabili notti fino alle quattro del mattino: il nostro consumo di caffeina potrebbe essere oggetto di ricerca. Al mio migliore amico Porto, che per me è come un fratello: sempre presente per sopportare le mie lagne ma anche mia principale fonte di risate e consigli. E naturalmente tutti i miei altri compagni di viaggio, come \emph{i Bibini} e \emph{i Pinguini}, che vorrei menzionare tutti se non fosse che sono le 3:40 del mattino mentre scrivo e faccio del mio meglio per non dimenticare nessuno.
    \switchcolumn*
    Second to last, as paradoxical as it may seem, I would like to express my gratitude to Quantum Mechanics, which has been the \emph{topic} that I enjoyed the most this year, both in its theoretical and experimental aspects. Sometimes hard\footnote{In this context, \emph{impossible}.} to understand but more often seductive and fascinating, it was one of the most appreciated discoveries I made during my university journey.
    \switchcolumn
    Per quanto possa sembrare paradossale, il mio penultimo ringraziamento è per la Meccanica Quantistica, la \emph{materia} che ho apprezzato di più quest'anno, sia nei suoi aspetti teorici che in quelli sperimentali. A volte difficile\footnote{In questo contesto, \emph{impossibile}.} da capire, ma più spesso seducente e affascinante, è stata una delle più belle scoperte che ho fatto durante il mio percorso universitario.
    \switchcolumn*
    Last but not least, my greatest gratitude goes to my family. Their support has been the invisible safety net beneath every step I have taken, even when they did not fully understand what I was doing or why I was still awake at absurd hours. Their love, patience and support have been the underlying solid ground that allowed me to make it this far.
    \switchcolumn
    Infine, rivolgo la mia più grande gratitudine ai miei genitori. Il loro supporto è stato la rete di sicurezza invisibile sotto ogni passo che ho fatto, anche quando non capivano cosa stessi affrontando o come mai fossi ancora sveglio a orari poco consoni. Il loro amore, la loro pazienza e il loro supporto sono stati il terreno solido che mi ha permesso di arrivare fino a questo punto.
\end{paracol}